---
title: "Pristine reef wanted dead or alive"
author: "Lepore ML, O’Dea A"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
# knitr ----
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  fig.align = "center",
  fig.asp = 0.618,
  fig.width = 6,
  out.width = "80%",
  cache = TRUE,
  echo = FALSE 
  )
# If you want to make sure the font size is consistent across all your figures,
# whenever you set out.width, you’ll also need to adjust fig.width to maintain
# the same ratio with your default out.width. For example, if your default
# fig.width is 6 and out.width is 0.7, when you set out.width = "50%" you’ll
# need to set fig.width to 4.3 (6 * 0.5 / 0.7).
```

# Motivate

### Shared context

Richard Condit built a useful R package. It has made the product of your work available to researchers in a shape they can analyse.

### Statement of the condition of the problem

But the potential of the R package underexploited. Researchers may be wasting time and producing inconsistent work, because they are building some tools from sratch in their own way, unaware if others built the same tool before or better.

### Statement of the consequences (Answer _So what?_)

Such efficiency leak can eventually impair not only the number of papers produced but also their  quality.

We may not be able to build an R package that is perfectly fit  for every reseracher, but we must try to understand what tools they need most, (practical and  cognitive), give them that, and support them along the process.

### Problem (condition + cost)

### Solution / main point / claim





R presentation

Use style to craft into

Shared context:



- make it easier to access, understand, and extend

- understand better what the community needs.


My role should be to ID those needs, and fulfill them through a system that has the R package st it's core but that depends on some other tools, practical and cognitive.



# Inform

## Role

Develop an [R package](http://ctfs.si.edu/Public/CTFSRPackage/) for use with repeated censuses of the CTFSForestGEO data.

### Goals:

- to update the R package

- to optimize the R package

- to establish a framework for future maintenance and expansion


## Some tasks

1. Produce and maintain a package sensu stricto

    - devtools; e.g. structure: [tidyverse](http://tidyverse.org/); phylosophy: [zoon](https://zoonproject.wordpress.com/); data: e.g. [babynames](https://github.com/hadley/babynames)

2. Set up a fully documented and active github

    - github repo, README.md, [pkgdown](http://hadley.github.io/pkgdown/index.html)

3. Fully document the functions and complete help files for each function. 

    - roxygen2,[pkgdown](http://hadley.github.io/pkgdown/index.html)

4. Optimize code for speed.

    - profvis, dplyr, c++

5. Integrate code written by other researchers into the package. 

    - github issues and pull requests

## Tasks most relevant to database managers

6. **Integrate the existing R functions to interact directly with output tables from the MySQL database**.

    - **dplyr to read directly database?**

7. Write new functions to help in the pre-processing and formatting of CTFS-ForestGEO data.

    - translate Condit's algorithm to R?

## Plan

1. Set up (1-2 months)

    - start with what we have
    
    - design what we want

2. Develop

    - **ID what users need**
    
    - build capacity (R workshop, e.g. replicate the one Owen organizes at STRI)
    
    - provide key functionality

    - support users, build trust, engage

3. Expand

    - incorporate users functionality

# Ask

## What else do we need?

[Stuart, Sean, Gabriel](https://mail.google.com/mail/u/0/#search/to%3Adaviess%40si.edu+has%3Aattachment/158c7ddcbb1b48db)

## How do you understand users?

## When is next census due?

## Is bookmarkable shiny state useful to database managers?

## Would people benefit from quarrying the database directly through dplyr?


## Note some inconsistent names

# Collaborate

- dplyr databases?

- visualization of data quality

- standarized preliminary analyses


# Toolsbox

<div class="notes">

xxxnotes; publish; allow pop-ups; add myslides.html?presentme=true (exit =false)

</div>

<div class="columns-2">

</div>

```{r}
# knitr::include_graphics("xxx")
```

# Todo

<div class="notes">



</div>

