---
title: "Pristine reef wanted dead or alive"
author: "Lepore ML, O’Dea A"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
# knitr ----
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  fig.align = "center",
  fig.asp = 0.618,
  fig.width = 6,
  out.width = "80%",
  cache = TRUE,
  echo = FALSE 
  )
# If you want to make sure the font size is consistent across all your figures,
# whenever you set out.width, you’ll also need to adjust fig.width to maintain
# the same ratio with your default out.width. For example, if your default
# fig.width is 6 and out.width is 0.7, when you set out.width = "50%" you’ll
# need to set fig.width to 4.3 (6 * 0.5 / 0.7).
```

# Inform

## Role

Develop an [R package](http://ctfs.si.edu/Public/CTFSRPackage/) for use with repeated censuses of the CTFSForestGEO data.

### Goals:

- to update the R package

- to optimize the R package

- to establish a framework for future maintenance and expansion


## Some tasks

1. Produce and maintain a package sensu stricto

    - devtools; e.g. structure: [tidyverse](http://tidyverse.org/); phylosophy: [zoon](https://zoonproject.wordpress.com/); data: e.g. [babynames](https://github.com/hadley/babynames)

2. Set up a fully documented and active github

    - github repo, README.md, [pkgdown](http://hadley.github.io/pkgdown/index.html)

3. Fully document the functions and complete help files for each function. 

    - roxygen2,[pkgdown](http://hadley.github.io/pkgdown/index.html)

4. Optimize code for speed.

    - profvis, dplyr, c++

5. Integrate code written by other researchers into the package. 

    - github issues and pull requests

## Tasks most relevant to database managers

6. **Integrate the existing R functions to interact directly with output tables from the MySQL database**.

    - **dplyr to read directly database?**

7. Write new functions to help in the pre-processing and formatting of CTFS-ForestGEO data.

    - translate Condit's algorithm to R?

## Plan

1. Set up (1-2 months)

    - start with what we have
    
    - design what we want

2. Develop

    - **ID what users need**
    
    - build capacity (R workshop, e.g. replicate the one Owen organizes at STRI)
    
    - provide key functionality

    - support users, build trust, engage

3. Expand

    - incorporate users functionality

# Ask

## What else do we need?

[Stuart, Sean, Gabriel](https://mail.google.com/mail/u/0/#search/to%3Adaviess%40si.edu+has%3Aattachment/158c7ddcbb1b48db)

## How do you understand users?

## When is next census due?

## Is bookmarkable shiny state useful to database managers?

## Would people benefit from quarrying the database directly through dplyr?


## Note some inconsistent names

# Collaborate

- dplyr databases?

- visualization of data quality

- standarized preliminary analyses


# Toolsbox

<div class="notes">

xxxnotes; publish; allow pop-ups; add myslides.html?presentme=true (exit =false)

</div>

<div class="columns-2">

</div>

```{r}
# knitr::include_graphics("xxx")
```

# Todo

<div class="notes">



</div>

