---
title: "Manageing and Sharing Data"
author: "Abby and Mauro  | slides at: rpubs.com/maurolepore/share"
date: "Presented on 2016-08-24 and reviewed on 2016-08-25"
output:
  ioslides_presentation:
    widescreen: false
---

```{r setup_slides, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA)
library(dplyr)
```

## Outline

1. Show a curated list of resources

2. Overview why and how using git(Hub)

3. How I use git(Hub) 


# 1. Show a curated list of resources

# 2. Overview why and how using git(Hub)

## Why using git(Hub)

- **to work safer.** It heps you work more safely so you can take more risks (not using git(Hub) is like free climbing)

- **to work neater** It helps you to organise your work better by using consistent structure and procedure

- **to work more effectively**. it helps you achieve goals easier by focusing on compleating small incremental steps (particular tasks) towards that goal

- **To show easier.**it provides a website for your project

- **To collaborate easier.**it provides a system for others to contribute to your work.




# 3. How I use git(Hub) 

- Install git, open an account in gitHub, link git(Hub) to RStudio

See section *Git and GitHub / Initial set up* [here](http://r-pkgs.had.co.nz/git.html)

- Create projects in R studio (R packages are projects). R Studio projects come with 

- Add some very basic components of the project

- Make an initial commit changes in R Studio

- Create a new repo in gitHub at https://github.com/new 

    + name the project exactly the same as its local version
    + click `create repository`
    + copy to clipboard what it says under **…or push an existing repository from the command line** which looks like this:

```
git remote add origin https://github.com/maurolepore/slides.git
git push -u origin master

```


- Push changes to git(Hub)



- Create a dev branch



Why hosting work on github? To share easy

If it is hosted on github, your work is easy to share.

E.g., to get the `coral` R package do this:

```{r, echo=TRUE, eval=FALSE}
install.packages("devtools")  # if you don't have it already

devtools::install_github("maurolepore/coral")
```



## Why hosting work on github? To control versions

### Version control allows you to:

- travel in time,

- take risks safely

- collaborate effectively


## Curated resources

1. Learn why using git and gitHub:  [webinar](https://www.rstudio.com/resources/webinars/collaboration-and-time-travel-version-control-with-git-github-and-rstudio/)

2. Learn how to use git and gitHub: [webinar](https://www.rstudio.com/resources/webinars/rstudio-essentials-webinar-series-managing-part-2/)

3. Get quick help on both git and gitHub: [gitHub help](https://help.github.com/)

4. Learn common actions using RStudio + git + gitHub: [chapter in R packages book](http://r-pkgs.had.co.nz/git.html)


5. Learn any topic to the greatest possible detail: [book Pro Git](https://git-scm.com/book/en/v2)



## To learn, pick a plan

### Quick but painful 

Gets you started quickly but hurts on your brain and comes with a lot of frustration. High risk of quitting.

- Use resource 3 and try git(Hub) as you watch. Use other resources as needed.

### Slowest but smoothest

Slowest but most effective. Your brain has sufficient time to digest the content and, before it forgets, the content is refreshed from a slightly different angle. The chances to quit are vey low and it can even be fun!

- Whatch resource 1 one day, then 2 the next day, then read 3 the next day AND try git(Hub) as you read. When you struggle, use resource 4 and 5.





## Manageing and Sharing Data Privately

`?install\_github(auth\_token = )`

> `auth\_token`: To install from a private repo, generate a personal access token (PAT) in https://github.com/settings/tokens and supply to this argument. This is safer than using a password because you can easily delete a PAT without affecting any others. Defaults to the GITHUB_PAT environment variable.





















