---
title: "Why and how to use git(Hub)"
author: "rpubs.com/maurolepore/git"
date: "2016-11-04"
output:
  ioslides_presentation:
    widescreen: false
---

```{r setup_slides, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA)
library(dplyr)
```

## Outline

1. Provide resources and suggestions to learn

2. Overview why and how using git(Hub)

3. How I use git(Hub) 


# 1. Provide resources and suggestions to learn

## A curated list of resources is this:

1. Learn why using git and gitHub:  [webinar](https://www.rstudio.com/resources/webinars/collaboration-and-time-travel-version-control-with-git-github-and-rstudio/)

2. Learn how to use git and gitHub: [webinar](https://www.rstudio.com/resources/webinars/rstudio-essentials-webinar-series-managing-part-2/)

3. Learn common actions using RStudio + git + gitHub: [chapter in R packages book](http://r-pkgs.had.co.nz/git.html)

4. Get quick help on both git and gitHub: [gitHub help](https://help.github.com/)


5. Learn any topic to the greatest possible detail: [book Pro Git](https://git-scm.com/book/en/v2)



## To learn, pick a plan that suits you

### Quick but painful 

Gets you started quickly but is boring and hurts your brain. Risks of frustration and quitting are high.

- Use resource 3 and try git(Hub) as you read. Use other resources as needed.

### Slowest but smoothest

Slowest but most effective. Your brain has sufficient time to digest the content and, before you forget, the content is refreshed from a slightly different angle. The chances to quit are vey low and it can even be fun!

- Whatch resource 1 one day, then watch 2 the next day, then read 3 the next day AND try git(Hub) as you read. When you struggle, use resource 4 and 5 as needed.

### Inbetween

- Watch resource 2 first, then use the **Quick but painful** plan.



# 2. Overview why and how using git(Hub)

## Why using git(Hub)

- **to work safer.** It heps you work more safely so you can take more risks (not using git(Hub) is like free climbing)

- **to work neater** It helps you to organise your work better by using consistent structure and procedure

- **to work more effectively**. it helps you achieve goals easier by focusing on compleating small incremental steps (particular tasks) towards that goal

- **To show easier.**it provides a website for your project

- **To collaborate easier.**it provides a system for others to contribute to your work.


## How to use git

- **From RStudio.** Start using git(Hub) from R studio, even if your project does not use R. You will gain a lot of important features in a very user-friendly way. Your will become familiar with how R Studio and git(hub) interact and, by the time your project uses R, you will be ready to go!

- **From the bash shell. ** Continue using git(Hub) from the shell to unleash the full power of git. For example, create tags and branches.




# 3. How I use git(Hub) 

- Install git, open an account in gitHub, link git(Hub) to RStudio

See section *Git and GitHub / Initial set up* [here](http://r-pkgs.had.co.nz/git.html)

- Create projects in R studio (R packages are projects). R Studio projects come with 

- Add some very basic components of the project

- Make an initial commit changes in R Studio

- Create a new repo in gitHub at https://github.com/new 

    + name the project exactly the same as its local version
    + click `create repository`
    + copy to clipboard what it says under **…or push an existing repository from the command line** which looks like this:

```
git remote add origin https://github.com/maurolepore/slides.git
git push -u origin master

```

- Paste what you just copied in a shell. Get the shell from R Studio (tools/shell). You need internet connection. You will need to provide the shell with your login and password to acess gitHub.

Done! From now on, new commits (in your master branch) can be pushed from RStudio to gitHub.



## Try git in RStudio

- Explore all you can do from RStudio

- Discuss alternative state of files, find help in 

- See history of commits, 

- interprate diff in each file and line

- revert a file or line





## More power from the shell

- tag

- branch







- Create a dev branch



Why hosting work on github? To share easy

If it is hosted on github, your work is easy to share.

E.g., to get the `coral` R package do this:

```{r, echo=TRUE, eval=FALSE}
install.packages("devtools")  # if you don't have it already

devtools::install_github("maurolepore/coral")
```



## Why hosting work on github? To control versions

### Version control allows you to:

- travel in time,

- take risks safely

- collaborate effectively

















