---
author: "Go to next slide"
output:
  ioslides_presentation:
    css: idigbio_styles.css
    smaller: yes
    widescreen: yes
  beamer_presentation: default
---

```{r removeSlideNumbers, include=FALSE}
# From https://github.com/rstudio/rmarkdown/issues/244
# To remove slide numbers to one of two things:

# One: Place this under the yamal header of the .Rmd file

# <style type="text/css">
# slides > slide:not(.nobackground):after {
#   content: '';
# }
# </style>



# Or, two: Place this in a separate mystyle.css file (remove <style> tag)

# slides > slide:not(.nobackground):after {
#   content: '';
# }

# For approach two, the yamal header should look like this:
# ---
# output:
#   ioslides_presentation:
    # css: mystyles.css
# ---
```

```{r setup, include=FALSE}
# knitr ----
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  fig.align = "center",
  fig.asp = 0.618,
  fig.width = 6,
  out.width = "75%",
  cache = TRUE,
  echo = FALSE 
  )
# If you want to make sure the font size is consistent across all your figures,
# whenever you set out.width, you’ll also need to adjust fig.width to maintain
# the same ratio with your default out.width. For example, if your default
# fig.width is 6 and out.width is 0.7, when you set out.width = "50%" you’ll
# need to set fig.width to 4.3 (6 * 0.5 / 0.7).
# plots
```

# Manage and Shar Data happily with GitHub and R

### Mauro Lepore

### **bookdown.org/maurolepore/idigbio**




### [wiki](https://goo.gl/p19EeC)

## Hit audience

The pit of success! Hadley

- pose question I want the audience to ask

Who ever heard of GitHub?
Who ever used GitHub?
Who regularly uses it?
Same for R packages?
--

Paleontologists...

In 2001, when I started my PhD, I had never heard GitHub, or R.

I was told it could help me, but I couldn't see how because the examples were so foreingn to my paleontological research.

The first time I tried R was horrible, the second too and many more.

Then I helped

...

Today, I am a software developer for ForestGEO, I write computer programs that help researchers to their research. ...

The point, I'm here to show you that GitHub can help you because it helped me.

   - not because I come from computer science and know every software detail
   - because of the opposite
   - because like you, I am a paleo ecologyst, and know how hard it is to manage and share data, and know only the surface of computer science, but that's enough to improve how we do science by a lot
   
   
--

Scientists work hard but their impact is limited because they share we manage and share data poorly. There is a tool that datascientists use, but few of us use it because the applications are so different to what we do that we struggle to see how we would benefit. But the connection is direct. What is that tool?
--

Managing and sharing data is difficult because xxx, AND dropbox and google docs help to some extent. BUT xxxx, THEREFORE we need more to do good science...
What is that tool?
--

There is a tool, commonly used among data scientists, that helps researchers too.  
...
what is that tool?


## Outline (message box)

Issue: A framework for to manage and share data

Audience: Researchers (paleontologists)

Problem: Managing and sharing data is difficult because collaborative research requires

    - different levels of access, 
    - editing simultaneously,
    - versioning,
    - citing versions,

So what? We don't get as much as we could from the data. Worse, data may be risked or unshared and research slowed.

Solution: GitHub (and R). dropbox and google docs are too poorly featured, and databases are too complicated. GitHub fills the gap:

    - simple,
    - free,
    - fully featured and dynamically improved,
    - popular.

Benefit: We can manage and share data more efficiently.




- message box?

# Opening

## From the know to the (maybe) unknown

# Action

## How to do it

- show examples. maybe slides, live programming may be clunky

### Give and get

COSTS TO YOU 

* 2' to open account

* 10' hello world

* ...

BENEFITS TO YOU

* website for your projects

* time travel

* discussion platform optionally bypass emails

* true collaboration

* privacy

* whats's here? when last changed? who changed it? Why? Oh, I want that version! https://speakerdeck.com/jennybc/happy-git-and-github-for-the-user


* diff https://speakerdeck.com/jennybc/happy-git-and-github-for-the-user

* metadata is the data https://speakerdeck.com/jennybc/happy-git-and-github-for-the-user

### Privacy


# ...

# Resolution

## Message

# Supplement

## Resources

Lots of links to everything. Here or **as I go**?

* GitHub hello world https://guides.github.com/activities/hello-world/

* Happy git and github for the useR http://happygitwithr.com/


* books by Hadley?

    * R Packages, chapter on git and github http://r-pkgs.had.co.nz/git.html

### presentation by Jenny Bryan

(see JB on bookmards toolbar)

[Happy Git and GitHub for the useR](https://speakerdeck.com/jennybc/happy-git-and-github-for-the-user)


## Invisible

<div class="notes">

SCOPE

- GitHub + R; but + * maybe OK
- No databases; but see dplyr
- from solo to ### collaborators
- (free)* software

* free private GitHub for science

STRUCTURE AND SYSTEMS

Data raw
    Document cleaning
Data private
    Document
        Private (private repo)
        Public (gh pages)
    Distribute (install_github)
        Private (GitHub privileges)
        Limited (hack: authorised guest token)
Data public
        Public (private repo)
        Public (gh pages)


1 project, public/private data
(1 local repo; 2 remotes)
  data-master push data
  data-private push data-private



TOOLS

dplyr


Document: metadata

Problems:
- Metadata unapdated (cookies in rice container, by Jenny Bryan)





Github & R: give structure

</div>






## Toolsbox

<div class="notes">

xxxnotes; publish; allow pop-ups; add myslides.html?presentme=true (exit =false)

</div>

<div class="columns-2">

</div>

```{r}
# knitr::include_graphics("xxx")
```

## Todo

<div class="notes">

xxx

</div>




