---
author: "Go to next slide"
output:
  ioslides_presentation:
    css: idigbio_styles.css
    smaller: yes
    widescreen: yes
  beamer_presentation: default
---

```{r removeSlideNumbers, include=FALSE}
# From https://github.com/rstudio/rmarkdown/issues/244
# To remove slide numbers to one of two things:

# One: Place this under the yamal header of the .Rmd file

# <style type="text/css">
# slides > slide:not(.nobackground):after {
#   content: '';
# }
# </style>



# Or, two: Place this in a separate mystyle.css file (remove <style> tag)

# slides > slide:not(.nobackground):after {
#   content: '';
# }

# For approach two, the yamal header should look like this:
# ---
# output:
#   ioslides_presentation:
    # css: mystyles.css
# ---
```

```{r setup, include=FALSE}
# knitr ----
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  fig.align = "center",
  fig.asp = 0.618,
  fig.width = 6,
  out.width = "75%",
  cache = TRUE,
  echo = FALSE 
  )
# If you want to make sure the font size is consistent across all your figures,
# whenever you set out.width, you’ll also need to adjust fig.width to maintain
# the same ratio with your default out.width. For example, if your default
# fig.width is 6 and out.width is 0.7, when you set out.width = "50%" you’ll
# need to set fig.width to 4.3 (6 * 0.5 / 0.7).
# plots
```

# Manage and Shar Data happily with GitHub and R

### Mauro Lepore

### **bookdown.org/maurolepore/idigbio**




### [wiki](https://goo.gl/p19EeC)

## Hit audience

The pit of success! Hadley

- pose question I want the audience to ask

## Outline (message box)

Issue: A framework for to manage and share data

Audience: Researchers (paleontologists)

Problem: Managing and sharing data is difficult because collaborative research requires

    - different levels of access, 
    - editing simultaneously,
    - versioning,
    - citing versions,

So what? We don't get as much as we could from the data. Worse, data may be risked or unshared and research slowed.

Solution: GitHub (and R). dropbox and google docs are too poorly featured, and databases are too complicated. GitHub fills the gap:

    - simple,
    - free,
    - fully featured and dynamically improved,
    - popular.

Benefit: We can manage and share data more efficiently.




- message box?

# Opening

## From the know to the (maybe) unknown

# Action

## How to do it

- show examples. maybe slides, live programming may be clunky

### Give and get

COSTS TO YOU 

* 2' to open account

* 10' hello world

* ...

BENEFITS TO YOU

* website for your projects

* time travel

* discussion platform optionally bypass emails

* true collaboration

* privacy

* whats's here? when last changed? who changed it? Why? Oh, I want that version! https://speakerdeck.com/jennybc/happy-git-and-github-for-the-user


* diff https://speakerdeck.com/jennybc/happy-git-and-github-for-the-user

* metadata is the data https://speakerdeck.com/jennybc/happy-git-and-github-for-the-user

### Privacy


# ...

# Resolution

## Message

# Supplement

## Resources

Lots of links to everything. Here or **as I go**?

* GitHub hello world https://guides.github.com/activities/hello-world/

* Happy git and github for the useR http://happygitwithr.com/


* books by Hadley?

    * R Packages, chapter on git and github http://r-pkgs.had.co.nz/git.html

### presentation by Jenny Bryan

(see JB on bookmards toolbar)

[Happy Git and GitHub for the useR](https://speakerdeck.com/jennybc/happy-git-and-github-for-the-user)


## Invisible

<div class="notes">

SCOPE

- GitHub + R; but + * maybe OK
- No databases; but see dplyr
- from solo to ### collaborators
- (free)* software

* free private GitHub for science

STRUCTURE AND SYSTEMS

Data raw
    Document cleaning
Data private
    Document
        Private (private repo)
        Public (gh pages)
    Distribute (install_github)
        Private (GitHub privileges)
        Limited (hack: authorised guest token)
Data public
        Public (private repo)
        Public (gh pages)


1 project, public/private data
(1 local repo; 2 remotes)
  data-master push data
  data-private push data-private



TOOLS

dplyr


Document: metadata

Problems:
- Metadata unapdated (cookies in rice container, by Jenny Bryan)





Github & R: give structure

</div>






## Toolsbox

<div class="notes">

xxxnotes; publish; allow pop-ups; add myslides.html?presentme=true (exit =false)

</div>

<div class="columns-2">

</div>

```{r}
# knitr::include_graphics("xxx")
```

## Todo

<div class="notes">

xxx

</div>




