---
title: "To manage/share data easily we can use R packages hosted on Github"
author: "maurolepore@gmail.com"
date: "2016-08-24"
output: ioslides_presentation
---

```{r setup_packages, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA)
library(dplyr)
library(coral)
```



## Outline

- Why using R packages?
- Why hosting R packages on github?
- What is an R package?
- How to build an R package?
- What if I don't use R?
- Examples of this approach



## Why using R packages?

To store and share data, functions and documentation.

E.g.:

```{r, eval=FALSE, echo=TRUE}
library(coral)
# Do this  |  To see documentation of:
?coral        # the package
?georef       # one dataset
?map          # one function
```



## Why using R packages? To share georef

```{r}
georef %>% select(region, time_interval, collected_from, sample_name)
```



## Why using R packages? To share map()

```{r}
map()
```







## What is an R package?

[See extra slide](https://github.com/maurolepore/tidy/blob/master/tidy-presentation/02packages.pdf)

A collection of folders and files with specific structure:

```
- data/:         data
- data-raw/:     code that cleans the raw data
- inst/extdata:  raw data
- man/:          documentation
- R/:            code of functions to share
- R/sysdata.rda: sata that the functions need but not the user
- R/utils.R:     code of handy functions not meant to be shared
- vignettes/:    docs mixing narrative and analyses (e.g. a paper)
- ...
```



## How to build an R package?

[Package writing in RStudio](https://www.rstudio.com/resources/webinars/rstudio-essentials-webinar-series-programming-part-3/)

[R packages (free online) book](http://r-pkgs.had.co.nz/) (Wickham, 2015).


Building a package is easy, really. For example:

- to create an R package, do this:

```{r, echo=TRUE, eval=FALSE}
devtools::create(path)
```

- to add data to the package, do this: 

```{r, echo=TRUE, eval=FALSE}
devtools::use_data(data)
```



## What if I don't use R?

Then, interact with the data using Shiny apps. They let you:

- [explore data](http://shiny.rstudio.com/gallery/datatables-demo.html)

- [summarize data](example http://shiny.rstudio.com/gallery/widgets.html)

- [download data](http://shiny.rstudio.com/gallery/file-download.html)

- [and much more](http://shiny.rstudio.com/gallery/), e.g. [a cool MDS](https://gallery.shinyapps.io/LDAelife/)

Learn more: http://shiny.rstudio.com/tutorial/



## Examples of this approach

- STRI's [ForestGEO](http://ctfs.si.edu/Public/CTFSRPackage/)

- [The Smithsonian Conservation Biology Institute](http://www.nsf.gov/awardsearch/showAward?AWD_ID=1458748)

- The [coral](https://maurolepore.github.io/coral/) project



## Discussion











