---
title: "Manageing and Sharing Data"
author: "Abby and Mauro  | slides at: rpubs.com/maurolepore/share"
date: "Presented on 2016-08-24 and reviewed on 2016-08-25"
output:
  ioslides_presentation:
    widescreen: false
---

```{r setup_slides, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA)
library(dplyr)
```



## Outline

1. Data should be tidy
2. To manage/share data easily we can use R packages
3. Scratchpads: What  has Abby learned so far

_15 min talk + 5 discusison  |  English (Q&A Spanish)_



# 1. Data should be tidy

```{r, child="tidy.Rmd"}

```




# 2. To manage/share data easily we can use R packages hosted on Github

```{r, child="packages.Rmd"}

```




# 3. Scratchpads: What I've learned so far

Go to [Abby's slides](https://docs.google.com/presentation/d/1pK3lb_rYjsHGSHaNnEFsDDxXUu8y9tyayp05dB7R3AM/edit?usp=sharing)



# Extras

## Manageing and Sharing Data Privately

`?install\_github(auth\_token = )`

> `auth\_token`: To install from a private repo, generate a personal access token (PAT) in https://github.com/settings/tokens and supply to this argument. This is safer than using a password because you can easily delete a PAT without affecting any others. Defaults to the GITHUB_PAT environment variable.





















