---
title: bookdown.com/forestgeoguest/workplan_ML
subtitle: "In construction"
author: "Mauro Lepore"
date: "2017-08-20"
output:
  ioslides_presentation:
    css: idigbio_styles.css
    smaller: yes
    widescreen: yes
  beamer_presentation: default
---

```{r removeSlideNumbers, include=FALSE}
# From https://github.com/rstudio/rmarkdown/issues/244
# To remove slide numbers to one of two things:

# One: Place this under the yamal header of the .Rmd file

# <style type="text/css">
# slides > slide:not(.nobackground):after {
#   content: '';
# }
# </style>



# Or, two: Place this in a separate mystyle.css file (remove <style> tag)

# slides > slide:not(.nobackground):after {
#   content: '';
# }

# For approach two, the yamal header should look like this:
# ---
# output:
#   ioslides_presentation:
    # css: mystyles.css
# ---
```

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  # fig.align = "center",
  fig.asp = 0.618,
  fig.width = 6,
  out.width = "100%",
  cache = TRUE,
  echo = FALSE 
  )
# If you want to make sure the font size is consistent across all your figures,
# whenever you set out.width, you’ll also need to adjust fig.width to maintain
# the same ratio with your default out.width. For example, if your default
# fig.width is 6 and out.width is 0.7, when you set out.width = "50%" you’ll
# need to set fig.width to 4.3 (6 * 0.5 / 0.7).
# plots
```



# Proposed work plan | Today-May 1, 2018

### __leporem\@si.edu__



## Outline

### Goal To discuss my workplan until May 1, 2018.

### Show tasks and proposed timeline (based on job description)

### Brake down tasks, discuss and edit timeline




# Tasks and timeline

# Tasks broken down | Discuss as we go over each task

# Summary

```{r}
caption <- "Summary of work plan for Sep'17-May'18."
tribble(
  
  ~task, ~priority, ~note,
  # -----------------------------------------------------------------------
  "xxx", "yyy", "zzz",
  "xxx", "yyy", "zzz",
  "xxx", "yyy", "zzz",
  "xxx", "yyy", "zzz",
  "xxx", "yyy", "zzz",
  "xxx", "yyy", "zzz",
  "xxx", "yyy", "zzz",
  "xxx", "yyy", "zzz"
  # -----------------------------------------------------------------------
)
```


## to incorporate




## Outline










Add lots of references of where I take my ideas from.



 # OVERARCHING GOALS (broken down)

CTFS-ForestGEO developed a package of analytical software in the R programming language. CTFS-ForestGEO seeks a person to:

1. update the package.
    1. curate: decide what functions are useful and worthwhile keeping. DONE
    1. document: improve help files.
        1. titles, description and details.
        1. argumets (reuse templates)
        1. authors
        1. group related functions in one place
        1. add examples
        1. build a website
    1. validate: validate inputs or warn and err accoridingly. This helps users
       to understand what went wrong.
    1. test: This helps developers to notice that something  went wrong so they 
       can promptly fix it.
        1. regression tests: check old and new versions of a function outut the
           same.
        1. test that functions do what they are expected to do.
        1. checks: devtools::check()
        1. __lintr__ to check for style
        1. __goodpractice::gp()__ to check for general good practices
        1. TravisCI to check package builds on all platforms.
    1. refactor: re-write code to improve it. E.g.
        1. names: make them more elocuent, consistent
        1. arguments: 
            1. name arguments more elocuently
            1. order arguments more sensibly (data, details)
            1. give arguments most common or safest defaults
        1. code: 
            1. improve style for better readability and consistency with modern
               style guides.
            1. improve safety
            1. reduce complexity
                1. reduce nestedness (mostly unnecessary _if_ blocks)
                1. remove the boilperate code of `for` loops by replaceing `for`
                   loops by functional programming approaches (most generally,
                   calls  to `lapply(X, FUN)`)
            1. improve clarity


1. optimize the package.
    1. Human brain(human brain): make useful code easier to find and use.
        1. naming functions better.
        1. grouping functions coherently.
        1. writting more complete and clearer documentation (goepen)
        1. facilitating access to documentation via websites
        1. announcing advances on twitter, workshops
        1. facilitating disussions on GitHub
        1. building capacity via code reviews, workshops, seminars, tutorials.
    1. Access to data:
        1. provide easy access to data for examples (__bci__, __bciex__).
        1. help the community to develop their own data packages.
        1. help the community to develop a friction-less workflow
    1. Code speed
        1. provide light data for examples and tests to run fast (__bciex__) 
        1. identify speed bottlenecks (__profviz__)
        1. replace slow by fast code. E.g. compare __base__ vs. __dplyr__ vs.
           __data.table__

1. establish a framework for future maintenance and expansion.
    1. build R packages _sensu stricto__
    1. create platforms for:
        1. github.com/forestgeo:
            1. hosting code repositories, packages and package websites
            1. discussing current issues, proposed changes, collaboration, etc.
            1. managing collaboration
        1. (rpubs.com/forestgeo): publishing R products publicly
        1. bookdown.com/forestgeoguest: publishing R products 
           with privileges control
        1. @forestgeo, #forestgeo, #forestr, etc.: announcing developments.



# GENERAL TRAITS

Clearly document help pages.
Write easy-to-read code.
Communicate and understand the needs of a broad range of researchers and students.



# TASKS

Catagories

(1) Produce and maintain a package sensu stricto

Done, and to be maintained over the year. Optimization and expantion is important but some updates come first (documentation, assertions, tests)

* Produce forestr: done.
* Produce other packages: as they become available: continuous.


(2) Set up a fully documented and active github (or other) repository for the package.

Done, and to be maintained over the year.

(3) Fully document the functions and complete help files for each function.

Work in progress. All priority functions to be documented  within the first year. Abundance module to be developed during September, and finished by the end of Septemer? 

An analysis of how many functions are needed for each function I want in a package is important. This would show how much effort really goes into developing each new function.

How many functions are there? How many in total if I count their dependencies? How many functions on average each new function forces me to develop (should decrease with increasing number of documented functions.) How many lines of code? How much is test coverage now? What is the recommendation?

(4) Optimize code for speed.

Towards the end of the first period

(5) Integrate code written by other researchers into the package.

As it becomes available. 

* __ttt__ by Sabrina  Russo, in a couple of months
* Ervan, by middle of the first term (January?)

(6) Integrate the existing R functions to interact directly with output tables from the MySQL database.

Needs clarification. What functions are those?

(7) Write new functions to help in the pre-processing and formatting of CTFS-ForestGEO data. 

Assuming these functions are the ones that Suzanne shared with me, i.e. those which purpose is to produce r-tables from database data (__rtbl__)

Probabbly after releasing some module of __forestr__, e.g. after releasing abundance. If so, I may work on __rtbl__ during October and hav it finished by the end of October.



# SKILLS

## Required skills: 

(1) Experience in building R packages, good understanding of S3 and S4 methods, and knowledge of generic R work as well as R environments. 

(2) Experience in git and github. 

(3) Experience in code debugging, profiling, and optimizing. 

(4) Strong attention to detail and organizational skills. 



## Desirable skills: 

(1) Experience with R Markdown and Shiny. 

(2) Experience with lower level languages, ideally C++. 

(3) Previous experience with forest ecology or plant dynamics.
